---
title: "Exponential Distribution and Simulated Sample Distribution"
author: "Omri Galai"
date: "12/18/2021"
output: pdf_document
---

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=50))
```
## Simulation
First let's simulate 1000 draws of 40 exponential random variables and store them in a data frame.
```{r}
set.seed(5000000)
nosim <- 1000
n <- 40
lambda <- 0.2
sim <- NULL
for(i in 1:nosim) sim = rbind(sim, rexp(n, lambda)) # 1000 runs of generating 40 r.vs
sim <- as.matrix(sim)
```
Now let's take a look at the dimensions of the data frame:
```{r}
dim(sim)
```
## Sample Mean and Variance versus Theretical Mean and Variance
Central limit theorem posits that the mean of sample averages approaches the normal dist., its mean equals population mean (1/lambda = 5, same as population sd) and its standard error  sigma/sqrt(n):
```{r}
means <- apply(sim, 1, mean) # taking the mean by rows
mean(means)
1/lambda # theoretical mean and s.d
sd(means)
print((1/lambda)/sqrt(n)) # theoretical standard error
```
As expected the simulation statistics are very close to computed values.
Visually, we can see in Figure 1 in the appendix that the population distribution and the sample distribution are centered around the same mean, while the distribution of sample means is less variable than the original distribution.
## Normality
See appendix for Figure 2 and code to generate it.
The histogram is centered at the mean, is mound shaped and is visibly similar to the normal density we have overlaid.
On the Q-Q plot. We observe an approximately linear relationship implying good fit normal distribution with more deviations toward the upper right corner representing a minimal right skewdness (based on [this blog post](https://towardsdatascience.com/q-q-plots-explained-5aa8495426c0).)
We can evaluate the same using skewness as well as kurtosis, which Wikipedia defines as a "measure of the "tailedness" of the probability distribution". The values for a normal distribution should be 0 and 3, respectively.
```{r}
library(moments)
skewness(means)
kurtosis(means)
```
Again, the data is mildly skewed to the right (skewness > 0) and has somewhat lighter tails than expected (kurtosis > 3). However, [rules](https://variation.com/wp-content/distribution_analyzer_help/hs139.htm) of [thumb](https://variation.com/wp-content/distribution_analyzer_help/hs113.htm) imply these are not sizable departure from normality. 
To sum up, we have seen that the distribution of 1000 samples of 40 observation each we generated approximates the expected statistical values very accurately and is very close to normal-shaped, as implied by the Central Line Theorem. Numeric evaluation exacerbates the visual impression on normality.

## Conclusion
In this report we showed that by taking a non-normal distribution, namely the exponential one, and sampling repeatedly from it a large number of times, we generated an approximately normal distribution of the *sample* means. After simulating 1000 iterations of n = 40 samples, we inspected the resulting distribution and saw its mean is very close to the population mean; its variability is smaller; and its means exhibit a good fit to a normal distribution, all in accordance with the central limit theorem. 
\newpage
# Appendix
```{r}
library(ggplot2)
seed <- 5000000
df <- data.frame(x = c(means, rexp(1000,0.2)), group = rep(c( "Sample Means", "Simulated Exponentials"), each = 1000))
g <- ggplot(df, aes(x, col = group)) + geom_histogram(binwidth = 0.1, alpha = 0.2) + geom_vline(xintercept = 5) + labs(title = "Histogram of Generated Exponentials and Sample Means")
g
```

```{r}
library(ggplot2)
library(gridExtra)
g2 <- g + stat_function(fun = dnorm, args = list(mean = mean(means), sd = sd(means)), colour = "red") # overlaying normal density
g2 <- g2 + labs(title = "Histogram of sample means (grey) and normal density (red)") + theme(title = element_text(size = 8)) # annotating
p <- ggplot(as.data.frame(means), aes(sample = means), xlab = "Sample means", ylab = "Density") 
p <- p + stat_qq(distribution = qnorm, colour = "black") + stat_qq_line(colour = "blue")
p <- p + labs(title = "Q-Q Plot", x = "Normal quantiles", y = "Sample quantiles") + theme(plot.title = element_text(size = 10.5)) # adding Q-Q plot
gridExtra::grid.arrange(g, p, nrow = 1, bottom = "Fig. 2") # arranging side by side

```