---
title: "Infernetial Analysis of Guinea Pig Tooth Growth Data"
author: "Omri Galai"
date: "12/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Infernetial Data Analysis Exercise
In this section we aim to derive conclusions about the ToothGrowth data in R package "datasets". We begin with attaching the package and looking at the data:
```{r}
library(datasets)
dim(ToothGrowth)
head(ToothGrowth)
```
To make sense of what this data represents we rely on the datasets documentation and see len represents a dependent variable/outcome - tooth length in guinea pigs, with supp[lement type] and dose being independent variables.
```{r}
summary(ToothGrowth)
xtabs(~supp+dose,data = ToothGrowth)
```
The 60 data points are split evenly between the Vitamin C and orange juice conditions, each containing 10 observations of 0.5, 1, and 2 mg dosages each.
Next we conduct t tests to compare the means in each dosage regimen:
```{r}
library(broom)
glance(t.test(ToothGrowth[ToothGrowth$dose == 1,]$len - ToothGrowth[ToothGrowth$dose == 0.5,]$len))
glance(t.test(ToothGrowth[ToothGrowth$dose == 2,]$len-ToothGrowth[ToothGrowth$dose == 1,]$len))
```
